---
output: github_document
always_allow_html: true
---

<!-- This file is used to create README.md
Note that the README.md document may need updating to change
'\<0.001' to '<0.001'. 
-->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

```

# reportRmd

<!-- badges: start -->
[![Lifecycle: Stable](https://img.shields.io/badge/lifecycle-stable-green.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![CRAN status](https://www.r-pkg.org/badges/version/reportRmd)](https://CRAN.R-project.org/package=reportRmd)
[![metacran downloads](https://cranlogs.r-pkg.org/badges/grand-total/reportRmd)](https://cran.r-project.org/package=reportRmd)
<!-- badges: end -->

The goal of reportRmd is to automate the reporting of clinical data in Rmarkdown environments. Functions include table one-style summary statistics, compilation of multiple univariate models, tidy output of multivariable models and side by side comparisons of univariate and multivariable models. Plotting functions include customisable survival curves, forest plots, and automated bivariate plots.

## Installation

Installing from CRAN:
``` r
install.packages('reportRmd')
```

You can install the development version of reportRmd from [GitHub](https://github.com/) with:


``` r
# install.packages("devtools")
devtools::install_github("biostatsPMH/reportRmd", ref="development")
```

## New Features

- Survival curves have been improved and now return ggplots
- Variable labels will now be automatically output in tables


## Documentation

[Online Documentation](https://biostatsPMH.github.io/reportRmd/)

## Examples

### Summary statistics by Sex

```{r example}
library(reportRmd)
data("pembrolizumab")
rm_covsum(data=pembrolizumab, maincov = 'sex',
covs=c('age','pdl1','change_ctdna_group'),
show.tests=TRUE)

```

### Using Variable Labels

```{r }
var_names <- data.frame(var=c("age","pdl1","change_ctdna_group"),
                          label=c('Age at study entry',
                                  'PD L1 percent',
                                  'ctDNA change from baseline to cycle 3'))

pembrolizumab <- set_labels(pembrolizumab,var_names)

rm_covsum(data=pembrolizumab, maincov = 'sex',
covs=c('age','pdl1','change_ctdna_group'))

```

### Multiple Univariate Regression Analyses


```{r example-2}
rm_uvsum(data=pembrolizumab, response='orr',
covs=c('age','pdl1','change_ctdna_group'))

```


### Tidy multivariable analysis

```{r}
glm_fit <- glm(orr~change_ctdna_group+pdl1+cohort,
               family='binomial',
               data = pembrolizumab)
rm_mvsum(glm_fit,showN=T)
```


### Combining univariate and multivariable models

```{r}
uvsumTable <- rm_uvsum(data=pembrolizumab, response='orr',
covs=c('age','sex','pdl1','change_ctdna_group'),tableOnly = TRUE)

glm_fit <- glm(orr~change_ctdna_group+pdl1,
               family='binomial',
               data = pembrolizumab)
mvsumTable <- rm_mvsum(glm_fit,tableOnly = TRUE)

rm_uv_mv(uvsumTable,mvsumTable)
```

### Simple survival summary table 
Shows events, median survival, survival rates at different times and the log rank test.
Does not allow for covariates or strata, just simple tests between groups
```{r}
 rm_survsum(data=pembrolizumab,time='os_time',status='os_status',
 group="cohort",survtimes=c(12,24),
# group="cohort",survtimes=seq(12,36,12),
# survtimesLbls=seq(1,3,1),
 survtimesLbls=c(1,2),
 survtimeunit='yr')

```

### Summarise Cumulative incidence

```{r}
library(survival)
data(pbc)
rm_cifsum(data=pbc,time='time',status='status',group=c('trt','sex'),
eventtimes=c(1825,3650),eventtimeunit='day')
```

### Plotting survival curves

```{r}
ggkmcif2(response = c('os_time','os_status'),
cov='cohort',
data=pembrolizumab)
```


### Plotting odds ratios

```{r, height=2}
require(ggplot2)
forestplot2(glm_fit)
```

### Plotting bivariate relationships

These plots are designed for quick inspection of many variables, not for publication.
```{r}
require(ggplot2)
plotuv(data=pembrolizumab, response='orr',
covs=c('age','cohort','pdl1','change_ctdna_group'))

```

