---
title: "Test Report.Rmd"
output: 
  pdf_document
  word_document
date: '2022-06-10'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# install.packages("devtools")
# devtools::install_github("biostatsPMH/reportRmd")

library(reportRmd)
```

# Data Sets

## pembrolizumab

Survival data with 15 variables on 94 patients
```{r}
?pembrolizumab
str(pembrolizumab)
```


## ctDNA
Repeated measures on participants over time tracking the change in tumour size and the ctDNA clearance status.
```{r}
?ctDNA
str(ctDNA)
```

## testData

This will NOT be submitted to CRAN - this is just for us to test with. There is a variable called badfactor that only has a single level, and here the cohort variable has a level that is just called cohort and another level called Cohort A. 

```{r}
?testData
str(testData)
```

# Testing Ideas

## Function Help File (ie?rm_covsum)

```{r}
?plotuv
```

The description is useful and doesn't have any types
The argument descriptions are clear
The see also section contains links to the relevant functions
Please suggest any examples you think would be useful

## Function Performance

```{r}
fit <- lm(l_size~age+sex+pdl1+baseline_ctdna,data=pembrolizumab)
rm_mvsum(fit)
```

Works with continuous, categorical (ordered) and binary response
Works with continuous response
Works with binary response
Works when there are factors with few observations
Works when there is missing data
Works when a variable name is contained in a factor name (testdata, badfactor)
P values output as expected
Output looks as expected knit in Word
Output looks as expected knit in pdf in pdf
For plots - legends look okay with multiple groups/single groups

Any marginal cases you can think of to make the function throw an error - or any problems you've experienced with reportRx

From Katherine:
```{r}
var.list <- c("age","sex","ph.ecog")
for (i in 1:3){
  f <- as.formula(paste0("Surv(time, status) ~ ph.karno +", var.list[i]))
  cox <- coxph(f, lung)
  rm_mvsum(cox, data = lung)
}
# FIXED


m1 <- crrRx(as.formula('time+status~sex'), lung)
rm_mvsum(model=m1, data = lung, showN = F)

# Error in mvsum(model = model, data = data, digits = digits, markup = FALSE,  : 
#   Supplied data frame does not have the correct number of non-missing rows.



```

