---
title: "Test Report.Rmd"
output: 
  pdf_document
  word_document
date: '2022-06-10'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# install.packages("devtools")
# devtools::install_github("biostatsPMH/reportRmd")

library(reportRmd)
```

# Data Sets

## pembrolizumab

```{r}
?pembrolizumab
str(pembrolizumab)
```


## ctDNA

```{r}
?ctDNA
str(ctDNA)
```

## testData

This will NOT be submitted to CRAN - this is just for us to test with 

```{r}
?testData
str(testData)
```

# Testing Ideas

## Function Help File (ie?rm_covsum)

```{r}
?plotuv
```

The description is useful and doesn't have any types
The argument descriptions are clear
The see also section contains links to the relevant functions
Please suggest any examples you think would be useful

## Function Performance

```{r}
fit <- lm(l_size~age+sex+pdl1+baseline_ctdna,data=pembrolizumab)
rm_mvsum(fit)
```

Works with continuous, categorical (ordered) and binary response
Works with continuous response
Works with binary response
Works when there are factors with few observations
Works when there is missing data
Works when a variable name is contained in a factor name
P values output as expected
Output looks as expected knit in Word
Output looks as expected knit in pdf in pdf



