---
title: "reportRmd"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
vignette: >
  %\VignetteIndexEntry{reportRmd}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

reportRmd is a package designed to facilitate the reporting of common statistical outputs easily in RMarkdown documents. The package supports pdf, html and word output without any changes to the body of the report. The main features are Table 1 style summaries, combining multiple univariate regression models into a single table, tidy multivariable model output and combining univariate and multivariable regressions into a single table. Single table summaries of median survival times and survival probabilities are also provided. A highly customisable survival curve function, based on ggplot2 can be used to create publication-quality plots. Visualisation plots are also available for bivariate relationships and logistic regression models. 

A word of caution. The reportRmd package is designed to facilitate statistical reporting and does not provide any checks regarding the suitability of the models fit. 

<!--
## Styles
Currently `html_pretty` supports three page themes, `cayman` (the default),
`tactile`, and `architect`. And there are also two syntax highlight styles:
`github` to mimic the syntax highlight on Github, and `vignette` that is used by
`html_vignette`. If no highlight parameter is given, the default style created
by Pandoc will be used.
-->

```{r setup, include=FALSE}
library(reportRmd)
knitr::opts_chunk$set(message = FALSE, warning = FALSE,dev="cairo_pdf")
```

# Summary statistics 

Basic summary statistics. Set IQR = TRUE for IQR instead of min/max. To show both use all.stats=TRUE


```{r}
rm_covsum(data=pembrolizumab, 
covs=c('age','sex'),IQR=TRUE)

```

This will produce summary statistics by Sex, and indicate which statistical test was used (`show.tests=TRUE`)

```{r}
rm_covsum(data=pembrolizumab, maincov = 'sex',
covs=c('age','pdl1','change_ctdna_group'),
show.tests=TRUE)

```

Group comparisons are non-parametric by default, specify `testcont='ANOVA'` for t-tests/ANOVA

```{r}
rm_covsum(data=pembrolizumab, maincov = 'sex',
covs=c('age','pdl1'),
testcont='ANOVA',
show.tests=TRUE)

```

The default is to indicate percentages by columns (ie percentages within columns add to 100)

```{r}
rm_covsum(data=pembrolizumab, maincov = 'sex',
covs=c('cohort'),
pvalue = FALSE)

```

But you can also specify to show by row instead

```{r}
rm_covsum(data=pembrolizumab, maincov = 'sex',
covs=c('cohort'),
pvalue = FALSE,
percentage='row')

```

# Univariate regression

Combining multiple univariate regression analyses into a single table

```{r }
rm_uvsum(data=pembrolizumab, response='orr',
covs=c('age','pdl1','change_ctdna_group'))

```

If the response is continuous linear regression is the default

```{r }
rm_uvsum(data=pembrolizumab, response='l_size',
covs=c('age','cohort'))

```

...unless two variables are specified and then survival analysis is run

```{r }
rm_uvsum(data=pembrolizumab, response=c('os_time','os_status'),
covs=c('age','pdl1','change_ctdna_group'))

```

Correlated observations can be handled using GEE

```{r}
 rm_uvsum(response = 'size_change',
 covs=c('time','ctdna_status'),
 gee=TRUE,
 id='id', corstr="exchangeable",
 family=gaussian("identity"),
 data=ctDNA,showN=TRUE)
```

If you want to check the underlying models, set `returnModels = TRUE`

```{r}
 rm_uvsum(response = 'orr',
 covs=c('age','sex','cohort'),
 data=pembrolizumab,returnModels = TRUE)
```

# Multivariable analysis

To create a nice display for multivariable models the model needs to first be fit.

```{r}
glm_fit <- glm(orr~change_ctdna_group+pdl1+age,
               family='binomial',
               data = pembrolizumab)
rm_mvsum(glm_fit, showN = TRUE)
```


# Combining univariate and multivariable models

To display both models in a single table run the rm_uvsum and rm_mvsum functions with `tableOnly=TRUE` and combine.

```{r}
uvsumTable <- rm_uvsum(data=pembrolizumab, response='orr',
covs=c('age','sex','pdl1','change_ctdna_group'),tableOnly = TRUE)

glm_fit <- glm(orr~change_ctdna_group+pdl1,
               family='binomial',
               data = pembrolizumab)
mvsumTable <- rm_mvsum(glm_fit, showN = TRUE,tableOnly = TRUE)

rm_uv_mv(uvsumTable,mvsumTable)
```

This can also be done with adjusted p-values, but when combined the raw p-values are dropped

```{r}
uvsumTable <- rm_uvsum(data=pembrolizumab, response='orr',
covs=c('age','sex','pdl1','change_ctdna_group'),tableOnly = TRUE,p.adjust='holm')

glm_fit <- glm(orr~change_ctdna_group+pdl1,
               family='binomial',
               data = pembrolizumab)
mvsumTable <- rm_mvsum(glm_fit,tableOnly = TRUE,p.adjust='holm')

rm_uv_mv(uvsumTable,mvsumTable)
```

Both raw and adjusted p-values still in the individual tables though

```{r}
outTable(mvsumTable)
```

# Simple Survival Summaries 

Displaying survival probabilities at different times by sex using Kaplan Meier estimates

```{r}
rm_survsum(data=pembrolizumab,time='os_time',status='os_status',
 group="sex",survtimes=seq(12,36,12),survtimeunit='months')

```

# Survival Times in Long Format

Displaying survival probabilities at different times by sex using Cox PH estimates

```{r}
rm_survtime(data=pembrolizumab,time='os_time',status='os_status',
 strata="sex",survtimes=c(12,24),survtimeunit='mo',type='PH')

```

Displaying survival probabilities at different times by sex, adjusting for age using Cox PH estimates

```{r}
rm_survtime(data=pembrolizumab,time='os_time',status='os_status', covs='age',
 strata="sex",survtimes=c(12,24),survtimeunit='mo',type='PH')

```


# Stratified Survival Summary
```{r}
rm_survdiff(data=pembrolizumab,time='os_time',status='os_status', 
            covs='sex',strata='cohort',digits=1)
```

# Plotting Functions

## Plotting bivariate relationships

These plots are designed for quick inspection of many variables, not for publication. This is the plotting version of rm_uvsum

```{r, , fig.width=7, fig.height=4}
plotuv(data=pembrolizumab, response='orr',
covs=c('age','cohort','pdl1','change_ctdna_group'))

```

## Plotting odds ratios

This will default to a log scale, but can be set to linear using `logscale=FALSE`
```{r, , fig.width=7, fig.height=2}
require(ggplot2)
glm_fit <- glm(orr~age+sex+change_ctdna_group+pdl1,
               family='binomial',
               data = pembrolizumab)
forestplot2(glm_fit)
```

## Plotting survival curves

```{r, fig.width=7,fig.height=5}
ggkmcif(response = c('os_time','os_status'),
cov='cohort',
data=pembrolizumab)
```


# PDF Output

For pdf to be correctly generate when using survival curves it is recommended that the cairo format be used. This can be specified with the following command in the setup code chunk:

`knitr::opts_chunk$set(message = FALSE, warning = FALSE,dev="cairo_pdf")`


# Data Sets

## pembrolizumab


Survival status and ctDNA levels for patients receiving pembrolizumab

A data frame with 94 rows and 15 variables:


- **id** Patient ID
- **age** Age at study entry
- **sex** Patient Sex
- **cohort** Study Cohort: A = Squamous cell carcinoma of soft pallate, B = Triple negative breast cancer, C = Ovarian, high grade serous, D = Melanoma, E = Other Solid Tumor
- **l_size** Target lesion size at baseline
- **pdl1** PD L1 percent
- **tmb** log of TMB
- **baseline_ctdna** Baseline ctDNA
- **change_ctdna_group** Did ctDNA increase or decrease from baseline to cycle 3
- **orr** Objective Response
- **cbr** Clinical Beneficial Response
- **os_status** Overall survival status, 0 = alive, 1 = deceased
- **os_time** Overall survival time in months
- **pfs_status** Progression free survival status, 0 = progression free, 1 = progressed
- **pfs_time** Progression free survival time in months



## ctDNA

Longitudinal changes in tumour size since baseline for patients by changes in ctDNA status (clearance, decrease or increase) since baseline.

A data frame with 270 rows and 5 variables:

- **id** Patient ID
- **cohort** Study Cohort: A = Squamous cell carcinoma of soft pallate, B = Triple negative breast cancer, C = Ovarian, high grade serous, D = Melanoma, E = Other Solid Tumor
- **ctdna_status** Change in ctDNA since baseline
- **time** Number of weeks on treatment
- **size_change** Percentage change in tumour measurement


