---
title: "reportRmd"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{reportRmd
  %\VignetteEngine{knitr::rmarkdown
  %\VignetteEncoding{UTF-8
---

```{r, include = FALSE
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup
library(reportRmd)
```

# Data Sets

## pembrolizumab


Survival status and ctDNA levels for patients receiving pembrolizumab

A data frame with 94 rows and 15 variables:


- **id** Patient ID
- **age** Age at study entry
- **sex** Patient Sex
- **cohort** Study Cohort: A = Squamous cell carcinoma of soft pallate, B = Triple negative breast cancer, C = Ovarian, high grade serous, D = Melanoma, E = Other Solid Tumor
- **l_size** Target lesion size at baseline
- **pdl1** PD L1 percent
- **tmb** log of TMB
- **baseline_ctdna** Baseline ctDNA
- **change_ctdna_group** Did ctDNA increase or decrease from baseline to cycle 3
- **orr** Objective Response
- **cbr** Clinical Beneficial Response
- **os_status** Overall survival status, 0 = alive, 1 = deceased
- **os_time** Overall survival time in months
- **pfs_status** Progression free survival status, 0 = progression free, 1 = progressed
- **pfs_time** Progression free survival time in months



## ctDNA

Longitudinal changes in tumour size since baseline for patients by changes in ctDNA status (clearance, decrease or increase) since baseline.

A data frame with 270 rows and 5 variables:

- **id** Patient ID
- **cohort** Study Cohort: A = Squamous cell carcinoma of soft pallate, B = Triple negative breast cancer, C = Ovarian, high grade serous, D = Melanoma, E = Other Solid Tumor
- **ctdna_status** Change in ctDNA since baseline
- **time** Number of weeks on treatment
- **size_change** Percentage change in tumour measurement

