---
title: "Testing reportRmd Labels"
format: 
  docx:
    fig-height: 2
    number-sections: false
    reference-doc: '../../../wordTemplate.docx'
editor: visual
---

Install the new version - this is on a separate branch, so that if you want to revert you the development or CRAN versions you can simply delete the reportRmd package folder and re-install.

```{r, echo=F,include=FALSE}
library(tidyverse)

```

```{r}
# if you don't have devtools installed:
#install("devtools")

# otherwise the variable label version is here:
#devtools::install_github("biostatsPMH/reportRmd", ref="varLabs")

# For the most recent stable version
# devtools::install_github("biostatsPMH/reportRmd", ref="development")
library(reportRmd)
```

# Labels & reportRmd

I've tried not to alter the basic workings of the package. The labels are added (and removed) from the variable attributes within data frames. **There is no longer an option for a default lookup table.**

# Worked Example

Get some descriptive stats for the ctDNA data that comes with the package. The `nicenames` argument is TRUE by default so underscores are replaced by spaces

```{r}
data(ctDNA)
rm_covsum(data=ctDNA,
          covs=c('cohort','ctdna_status','size_change'))

```

If we have a lookup table of variable names and labels that we imported from a data dictionary we can set the variable labels for the data frame and these will be used in the `rm_` functions

```{r}

ctDNA_names <- data.frame(var=names(ctDNA),
                          label=c('Patient ID',
                                  'Study Cohort',
                                  'Change in ctDNA since baseline',
                                  'Number of weeks on treatment',
                                  'Percentage change in tumour measurement'))
ctDNA <- set_labels(ctDNA,ctDNA_names)

rm_covsum(data=ctDNA,
          covs=c('cohort','ctdna_status','size_change'))

```

We can change individual labels with the `set_var_labels` command

```{r}
ctDNA <- set_var_labels(ctDNA,
                        cohort="A new cohort label")
rm_covsum(data=ctDNA,
          covs=c('cohort','ctdna_status','size_change'))

```

We can extract the variable labels to a data frame

```{r}
var_labels <- extract_labels(ctDNA)
var_labels
```

Or clear them all

```{r}
ctDNA <- clear_labels(ctDNA)
```
