---
title: "reportRmd Refactoring Validation"
format: html
---

```{r setup}
library(reportRmd)
library(survival)
```

## TEST Data

```{r create-data}
set.seed(42)
n <- 100

test_data <- data.frame(
  # Continuous: complete
  age = round(rnorm(n, mean = 55, sd = 12), 1),
  # Continuous: ~15% missing
  biomarker = ifelse(runif(n) < 0.15, NA,
                     round(rlnorm(n, meanlog = 2, sdlog = 0.8), 2)),
  # Categorical: 3 levels, complete
  stage = factor(sample(c("I", "II", "III"), n, replace = TRUE,
                        prob = c(0.3, 0.4, 0.3))),
  # Categorical: 4 levels, ~10% missing
  treatment = factor(ifelse(runif(n) < 0.10, NA,
                            sample(c("Placebo", "Low", "Medium", "High"), n,
                                   replace = TRUE,
                                   prob = c(0.25, 0.25, 0.25, 0.25)))),
  # Binary grouping variable
  sex = factor(sample(c("Male", "Female"), n, replace = TRUE))
)

# Add survival outcomes (for uvsum/mvsum survival tests)
test_data$os_time <- round(rexp(n, rate = 0.05), 1)
test_data$os_status <- rbinom(n, 1, prob = 0.4)

# Add a binary outcome for logistic tests
test_data$response <- rbinom(n, 1, prob = 0.3)

str(test_data)
summary(test_data)
```


# 1. rm_covsum — covsum_prepare, covsum_cramers_v, fmt_pct

## Basic (no grouping)

```{r covsum-basic}
rm_covsum(data = test_data,
          covs = c("age", "biomarker", "stage", "treatment"))

```

Same - with tidyselect
```{r}
test_data |> rm_covsum(
          covs = c(age, biomarker, stage, treatment))

```

## Grouped by sex

```{r covsum-grouped}
rm_covsum(data = test_data,
          covs = c("age", "biomarker", "stage", "treatment"),
          maincov = "sex")
```

## With effect sizes and test names (exercises covsum_cramers_v)

```{r covsum-effsize}
rm_covsum(data = test_data,
          covs = c("age", "biomarker", "stage", "treatment"),
          maincov = "sex",
          effSize = TRUE, show.tests = TRUE)
```

## Row percentages

```{r covsum-row-pct}
rm_covsum(data = test_data,
          covs = c("stage", "treatment"),
          maincov = "sex",
          percentage = "row")
```

## Custom digits

```{r covsum-digits}
rm_covsum(data = test_data,
          covs = c("age", "biomarker", "stage"),
          maincov = "sex",
          digits = 3, digits.cat = 1)
```

## tableOnly (returns data.frame)

```{r covsum-tableonly}
tab <- rm_covsum(data = test_data,
                 covs = c("age", "biomarker", "stage", "treatment"),
                 maincov = "sex",
                 tableOnly = TRUE)
str(tab)
```


# 2. outTable — header_above, round_numeric_cols

## header_above with grouped covsum

```{r outtable-header-above}
tab <- rm_covsum(data = test_data,
                 covs = c("age", "biomarker", "stage"),
                 maincov = "sex",
                 tableOnly = TRUE)

outTable(tab,
         header_above = c(" " = 1, " " = 1, "Sex" = 2, " " = 1))
```

## outTable with numeric rounding

```{r outtable-rounding}
num_tab <- data.frame(
  Variable = c("A", "B", "C"),
  Mean = c(1.23456, 7.89012, 3.45678),
  SD = c(0.12345, 0.67890, 0.23456)
)
outTable(num_tab, digits = 2)
```


# 3. rm_compactsum — apply_cat_test, fmt_pct

## Basic compact table

```{r compactsum-basic}
rm_compactsum(data = test_data,
              xvars = c("age", "biomarker", "stage", "treatment"))
```

## Grouped compact table (exercises fmt_pct in grouped path)

```{r compactsum-grouped}
rm_compactsum(data = test_data,
              xvars = c("age", "biomarker", "stage", "treatment"),
              grp = "sex")
```

## With p-values, effect sizes, test names (exercises apply_cat_test)

```{r compactsum-pvalue}
rm_compactsum(data = test_data,
              xvars = c("age", "biomarker", "stage", "treatment"),
              grp = "sex",
              pvalue = TRUE, effSize = TRUE, show.tests = TRUE)
```

## Row percentages (exercises fmt_pct row-percentage path)

```{r compactsum-row-pct}
rm_compactsum(data = test_data,
              xvars = c("stage", "treatment"),
              grp = "sex",
              percentage = "row")
```

## Custom category digits (exercises fmt_pct with digits.cat > 0)

```{r compactsum-digits}
rm_compactsum(data = test_data,
              xvars = c("stage", "treatment"),
              grp = "sex",
              digits.cat = 1)
```


# 4. rm_uvsum — autoreg_simple_fit, format_bold_pvalues, gp_from_drop1

## Linear model

```{r uvsum-linear}
rm_uvsum(response = "age",
         covs = c("biomarker", "stage", "treatment", "sex"),
         data = test_data)
```

## Logistic model

```{r uvsum-logistic}
rm_uvsum(response = "response",
         covs = c("age", "biomarker", "stage", "treatment", "sex"),
         data = test_data)
```

## Cox model

```{r uvsum-cox}
rm_uvsum(response = c("os_time", "os_status"),
         covs = c("age", "biomarker", "stage", "treatment", "sex"),
         data = test_data)
```

## With p-value adjustment

```{r uvsum-padj}
rm_uvsum(response = "age",
         covs = c("biomarker", "stage", "treatment", "sex"),
         data = test_data,
         p.adjust = "bonferroni")
```

## Global p-values only

```{r uvsum-global-p}
rm_uvsum(response = "age",
         covs = c("biomarker", "stage", "treatment", "sex"),
         data = test_data,
         whichp = "global")
```

## Both level and global p-values

```{r uvsum-both-p}
rm_uvsum(response = "age",
         covs = c("biomarker", "stage", "treatment", "sex"),
         data = test_data,
         whichp = "both")
```


# 5. rm_mvsum — coeffSum, gp, get_model_data, format_bold_pvalues

## Linear model

```{r mvsum-lm}
fit_lm <- lm(age ~ biomarker + stage + treatment + sex, data = test_data)
rm_mvsum(fit_lm)
```

## Logistic model

```{r mvsum-glm}
fit_glm <- glm(response ~ age + biomarker + stage + sex,
               data = test_data, family = "binomial")
rm_mvsum(fit_glm)
```

## Cox model

```{r mvsum-cox}
fit_cox <- coxph(Surv(os_time, os_status) ~ age + stage + sex,
                 data = test_data)
rm_mvsum(fit_cox)
```

## With VIF

```{r mvsum-vif}
rm_mvsum(fit_lm, vif = TRUE)
```

## With unadjusted estimates

```{r mvsum-unadjusted}
rm_mvsum(fit_lm, include_unadjusted = TRUE)
```

## Global p-values

```{r mvsum-global-p}
rm_mvsum(fit_lm, whichp = "both")
```

## With p-value adjustment

```{r mvsum-padj}
rm_mvsum(fit_lm, p.adjust = "holm")
```


# 6. pstprn / psthr — compact parameter

These helpers were refactored with a `compact` flag. They are exercised
indirectly through the model summaries above, but let's verify directly.

```{r pstprn-psthr}
# Normal formatting (spaces, commas)
cat("pstprn:", reportRmd:::pstprn(c(1.5, 0.8, 2.3)), "\n")
cat("psthr:", reportRmd:::psthr(c(1.5, 0.8, 2.3), y = 2), "\n")

# Compact formatting (no spaces)
cat("pstprn0:", reportRmd:::pstprn0(c(1.5, 0.8, 2.3)), "\n")
cat("psthr0:", reportRmd:::psthr0(c(1.5, 0.8, 2.3), digits = 2), "\n")
```


# 7. Survival plots — format_ci, append_hr_labels

```{r ggkmcif-km, fig.width=8, fig.height=6}
ggkmcif2(response = c("os_time", "os_status"),
         cov = "sex",
         data = test_data)
```

```{r ggkmcif-hr, fig.width=8, fig.height=6}
ggkmcif2(response = c("os_time", "os_status"),
         cov = "sex",
         data = test_data,
         HR = TRUE, HR_pval = TRUE)
```

```{r ggkmcif-stage, fig.width=8, fig.height=6}
ggkmcif2(response = c("os_time", "os_status"),
         cov = "stage",
         data = test_data,
         HR = TRUE, HR_pval = TRUE)
```


# 8. nestTable — round_numeric_cols

```{r nesttable}
m1 <- rm_mvsum(fit_lm, tableOnly = TRUE)
m1$Model <- "Linear"

fit_lm2 <- lm(biomarker ~ age + stage + sex, data = test_data)
m2 <- rm_mvsum(fit_lm2, tableOnly = TRUE)
m2$Model <- "Biomarker"

nestTable(rbind(m1, m2), head_col = "Model", to_col = "Covariate")
```


# 9. gp methods — gp_from_drop1, gp_crr_lrt

The gp generics are exercised through rm_mvsum and rm_uvsum above.
Let's also call them directly to verify the shared helpers.

```{r gp-direct}
# gp() is now a properly registered S3 generic — dispatch works correctly.

# gp.default (lm) — uses gp_from_drop1
gp_lm <- reportRmd::gp(fit_lm)
cat("gp.default (lm):\n")
print(gp_lm)

# gp.glm — uses gp_from_drop1 with test = "LRT"
gp_glm <- reportRmd::gp(fit_glm)
cat("\ngp.glm:\n")
print(gp_glm)

# gp.coxph — uses gp_from_drop1 with custom extract_pvals
gp_cox <- reportRmd::gp(fit_cox)
cat("\ngp.coxph:\n")
print(gp_cox)
```


# 10. Edge cases

## All missing in one group for a categorical variable

```{r edge-all-missing}
edge_data <- test_data
edge_data$treatment[edge_data$sex == "Male"] <- NA
rm_compactsum(data = edge_data,
              xvars = c("treatment"),
              grp = "sex")
```

## Single-level factor after dropping missing

```{r edge-single-level}
edge_data2 <- test_data
edge_data2$stage <- factor("I", levels = c("I", "II", "III"))
rm_covsum(data = edge_data2,
          covs = c("age", "stage"),
          maincov = "sex")
```

## Unformatted p-values

```{r edge-unformatted-p}
rm_uvsum(response = "age",
         covs = c("stage", "sex"),
         data = test_data,
         unformattedp = TRUE, tableOnly = TRUE)
```


# Cluster Data
```{r}
set.seed(123)

# 60 clusters: 40 observed twice (80 obs) + 20 observed once (20 obs) = 100 obs
n_twice <- 40
n_once <- 20
n_clusters <- n_twice + n_once

ids_twice <- 1:n_twice
ids_once <- (n_twice + 1):(n_twice + n_once)

# Build id vector
id <- c(rep(ids_twice, each = 2), ids_once)
n_obs <- length(id)

# Cluster-level covariates
age_cluster <- round(rnorm(n_clusters, mean = 60, sd = 12))
sex_cluster <- sample(c("Female", "Male"), n_clusters, replace = TRUE)

# Expand to observation level
age <- age_cluster[id]
sex <- factor(sex_cluster[id])

# Time since baseline: 0 for first obs, positive for second
time <- numeric(n_obs)
time[seq(2, n_twice * 2, by = 2)] <- round(runif(n_twice, min = 3, max = 24))

# Continuous outcome (correlated within cluster)
cluster_effect <- rnorm(n_clusters, sd = 5)
y_cont <- round(50 + 0.3 * age - 2 * as.numeric(sex == "Male") + 3 * time / 12 +
                 cluster_effect[id] + rnorm(n_obs, sd = 4), 1)

# Binary outcome
lp <- -2 + 0.03 * age + 0.5 * as.numeric(sex == "Male") + 0.05 * time + cluster_effect[id] * 0.1
prob <- plogis(lp)
y_bin <- rbinom(n_obs, 1, prob)

# Status variable (cluster-level event indicator)
status_cluster <- rbinom(n_clusters, 1, 0.6)
status <- status_cluster[id]

test_cluster_data <- data.frame(
  id = factor(id),
  time_from_baseline = time,
  status = status,
  y_binary = y_bin,
  y_continuous = y_cont,
  age = age,
  sex = sex
)

model <- geepack::geeglm(y_binary~age+sex,
                 data = test_cluster_data,
                 family = "binomial",
                 id = test_cluster_data$id)
summary(model)
rm_mvsum(model)

# re-order
test_cluster_data <- test_cluster_data |> 
  dplyr::arrange(y_continuous)
model <- geepack::geeglm(y_binary~age+sex,
                 data = test_cluster_data,
                 family = "binomial",
                 id = test_cluster_data$id)
summary(model)
#rm_mvsum(model) # should fail, because of mis-ordered clusters

# order properly
test_cluster_data <- test_cluster_data |> 
  dplyr::arrange(id)
model <- geepack::geeglm(y_binary~age+sex,
                 data = test_cluster_data,
                 family = "binomial",
                 id = test_cluster_data$id)
summary(model)

```


# Testing the Event Counts for Clustered models

## TEST 1: GEE - continuous outcome

```{r}
gee_cont <- geepack::geeglm(y_continuous ~ age + sex + time_from_baseline,
                              data = test_cluster_data, id = id,
                              family = gaussian, corstr = "exchangeable")
summary(gee_cont)
rm_mvsum(gee_cont, showN = TRUE, vif = FALSE)

```


## TEST 2: GEE - binary outcome


```{r}
gee_bin <- geepack::geeglm(y_binary ~ age + sex + time_from_baseline,
                             data = test_cluster_data, id = id,
                             family = binomial, corstr = "exchangeable")
summary(gee_bin)
rm_mvsum(gee_bin, showN = TRUE, showEvent = TRUE, vif = FALSE,tableOnly = F)

```


## TEST 3: lmer - continuous outcome


```{r}
lmer1 <- lme4::lmer(y_continuous ~ age + sex + time_from_baseline + (1|id),
                     data = test_cluster_data)
summary(lmer1)
rm_mvsum(lmer1, showN = TRUE, vif = FALSE)

model <- lme4::lmer(y_continuous ~ age + sex + time_from_baseline + (1|id),
                     data = test_cluster_data)
summary(lmer1)

```


## TEST 4: glmer - binary outcome


```{r}
glmer1 <- lme4::glmer(y_binary ~ age + sex + time_from_baseline + (1|id),
                       data = test_cluster_data, family = binomial)
summary(glmer1)
rm_mvsum(glmer1, showN = TRUE, showEvent = TRUE, vif = FALSE)
```



## TEST 5: coxph with cluster id


```{r}
cox1 <- survival::coxph(survival::Surv(time_from_baseline + 0.1, status) ~ age + sex,
                         data = test_cluster_data, id = id)
summary(cox1)
rm_mvsum(cox1, showN = TRUE, showEvent = TRUE, vif = FALSE)

```


## TEST 6: Regular glm (no clustering) - for comparison




```{r}
glm1 <- glm(y_binary ~ age + sex + time_from_baseline,
             data = test_cluster_data, family = binomial)
summary(glm1)
rm_mvsum(glm1, showN = TRUE, showEvent = TRUE, vif = FALSE,tableOnly = T)

```

# Test of cox models with formula

```{r}
time <- "time_from_baseline"
status <- "status"
group <- "sex"
surv_formula <- as.formula(
    paste0("Surv(", time, ", ", status, ") ~ ", group)
  )
  
# Cox fit for HR
coxfit <- coxph(surv_formula, data = test_cluster_data)
summary(coxfit)
rm_mvsum(coxfit)
```

